Galaxy Zoo: Morphological Classifications for 120,000 Galaxies in HST Legacy Imaging

### 2016-05-29 writeup/gz_hubble_data.tex
Text changes from KLM. Closes #169.

% Note: 250-word limit
We present the data release paper for the Galaxy Zoo: Hubble (GZH)
project.\footnote{This publication has been made possible by the participation
of more than 200,000~volunteers in the Galaxy Zoo project. Their contributions
are individually acknowledged at
\url{http://authors.galaxyzoo.org/authors.html}.} This is the third phase in a
large effort to measure reliable, detailed morphologies of galaxies by using
crowdsourced visual classifications of color composite images. Images in GZH
were selected from various \textit{Hubble Space Telescope} Legacy programs
using the Advanced Camera for Surveys, with filters that probe the rest-frame
optical emission from galaxies out to $z\sim1$.  The bulk of the sample is
selected to have $m_{I814W}<23.5$ across AEGIS, COSMOS, GEMS, and 2-epoch
exposures in GOODS-N and GOODS-S. Full-depth 5-epoch images from the GOODS
fields are also included and have $m_{I814W}<26.8$. The median redshift of the
combined samples is $\langle z\rangle=0.9\pm0.6$, with a tail extending out to
$z\simeq4$. The GZH morphological data include measurements of both bulge- and
disk-dominated galaxies, details on spiral disk structure that relate to the
Hubble type, bar identification, and numerous measurements of clump
identification and geometry.  This paper also describes a new method for
calibrating morphologies for galaxies of different luminosities and at
different redshifts by using artificially-redshifted galaxy images as a
baseline. The GZH catalog contains both raw and calibrated morphological vote
fractions for $119,849$~galaxies, providing the largest dataset to date
suitable for large-scale studies of galaxy evolution out to $z\sim1$. 
% Number = hubble + goods_shallow + stripe82_coadd
\end{abstract}

@@

@@

\item The Galaxy Evolution from Morphologies and SEDs
    \citep[GEMS;][]{rix04,cal08} survey was also centered on the Chandra Deep
    Field-South. The GEMS data covered $\sim800$~arcmin$^2$, completely surrounding
    the area covered by GOODS-S. Images from ACS in GEMS had 1~orbit per pointing
    for a total of 63~pointings. The exposure times were 2160 and 2286~seconds in
    \Vband{} and \zband{}, respectively. The image resolution had a pixel scale
    of 0.03~\arcsec/pixel. The $5\sigma$ limiting magnitude for source
    detection was 25.7~AB in \Vband{} and 24.2~AB in \zband.

@@

The project includes images from the SDSS \citep{yor00,str02}, in particular
those from Data~Release~7 \citep{aba09}. A small number of images were used in
the creation of simulated \hst{} images, described in Sections
\ref{ssec:ferengi} and \ref{sec:debiasing} below; the majority of SDSS images
used in GZH were from Stripe~82. The Stripe~82 samples in GZH include
single-epoch and co-added images. The single-epoch images provide a local
sample for comparison to higher-redshift galaxies (including the new
measurments of clumpy structure), while the co-added images allow for analysis
of morphological properties as a function of image depth.

@@

\begin{figure*}
\center
\includegraphics[width=160mm]{figures/gzh_interface.png}
\caption{Screenshot of the GZH interface (\url{http:/zoo3.galaxyzoo.org}) at
the beginning of a classification, with the classifier ready to select an answer for
the first question in the decision tree.\label{fig:interface}}
\end{figure*}

@@

Few galaxies in the sample have sufficiently high corrections to completely
change them from being confidently ``smooth'' to ``featured'' following the
correction for redshift bias (Figure~\ref{fig:debiased_corrections}). As a
check, highly boosted galaxies are compared to the expert classifications in
CANDELS \citep{kar15}.  There are only eight galaxies that were strongly
boosted (\ffeatures$<0.2$ and \fbest$>0.5$) in GZH and also appear in the
CANDELS expert sample. Of those eight galaxies, \citet{kar15} classify 5 as
spheroids, 2 as disks, and 1 as irregular/disky. The \fbest{} values for GZH
are all between 0.5 and 0.6, making them intermediate disk candidates. One of
the spheroids has a hint of extended emission to the south that may have been
missed by the CANDELS experts, but almost all images are dim and with very low
surface brightnesses in the ACS imaging, making positive identification
challenging. Since the surveys use different rest-frame filters and the overlap
sample is tiny, though, detailed comparisons between the overall morphologies
are highly difficult.

@@

As with the HST images, each \ferengi{} subject had a varying number of classifiers
answering the various questions in the hierarchical decision tree. Every classifier
answers the first question, {\it ``Is the galaxy smooth and rounded, with no
sign of a disk?''}; as such the vote fractions \fsmooth, \ffeatures, and
\fartifact{} are all computed with the minimum statistical error for any
question, with between 40 and 140~responses (see Section~\ref{sec:interface}).
The number of participants answering subsequent questions, however, is always equal
to or less than the number who answered the preceding question. The average number
of responses per task for fourth- or fifth-tier questions (such as spiral arm
structure; Tasks 12--14) is only $4\pm4$ for the \ferengi{} sample. While this
distribution is strongly bimodal (reflecting the true morphologies of
selected galaxies), the very low absolute numbers of votes introduce very high
variance when attempting to calculate a statistical correction.

@@

Another existing study using GZH to select barred galaxies was presented in
\citet{che15}. This study used a slightly different bar selection method to
demonstrate that AGN hosts show no statistically significant enhancement in bar
fraction or average bar likelihood compared to closely-matched inactive
galaxies. We note that their technique matched test and control galaxies over
the same redshift ranges, which minimizes the possible impact of redshift bias.

@@

COSMOS galaxies have multiple published datasets automatically classifying
morphology, all using a variation of non-parametric measurements. \citet{cas07}
used a combination of concentration ($C$), asymmetry ($A$), $G$, and $M_{20}$
\citep{cas05} make empirical divisions based on several hundred training images
to assign galaxy morphology. A similar method was employed by \citet{tas11},
using the same non-parametric indices but with a different method of
calculating the Petrosian radius and total light profile, and using a
nearest-neighbors method to label morphologies. \citet[][ZEST]{sca07} used $C$,
$A$, $G$, $M_{20}$, the galaxy ellipticity ($\epsilon$), and \sersic{} index
($n$) to quantify morphology; a principal component analysis is used to assign
galaxies to categories. All three methods use discrete morphological categories
of ellipticals, spirals, and peculiar/irregulars. We use the same categories
with the exception of assigning S0/bulge-dominated disks (type~2.0) from ZEST
into the elliptical category.

@@

Galaxies from the three COSMOS surveys all have morphologies that are
reasonably well-predicted by the GZH based on \fbest. The relation is strongest
for early-type galaxies, since the comparable fractions of irregulars and
spirals shows that clumpy/asymmetric morphologies are clearly also being
captured by the \fbest{} parameter. The small dropoff in the fraction of
early-types for the Cassata and ZEST samples at \fbest$<0.1$ is somewhat
puzzling, but could potentially be due to the presence of a small
population of extremely compact objects (stars or quasars) that would be
classified as artifacts in GZH, explaining the dip in \fbest.

We also emphasize that the various COSMOS catalogs are not consistent; for
example, only 64\% of the galaxies in \citet{tas11} have the same morphological
class as \citet{cas07}. The bulk of these are galaxies classified as ``spiral''
using one technique and ``irregular'' using another (\citealt{cas07} is more
likely to classify galaxies as irregular).  A detailed comparison of the
external catalogs is outside the scope of this paper, but we emphasize only
that it is \emph{impossible} for any catalog to be consistent with all three.
We interpret the broad agreements between early-types, spirals, and irregulars
for all three samples as a validation of the crowdsourced visual
classifications provided by GZH.

\begin{figure*}
\center
\includegraphics[width=1.0\textwidth]{figures/comparisons_odd.pdf}
\caption{Same as Figure~\ref{fig:comparisons_features}, but comparing external
morphologies to \fodd{} in GZH. Measurement of \fodd{} excludes galaxies
labeled as ``odd'' due to lenses or dust lanes ($<1\%$ of the total), which are GZH
categories not expected to correlate with larger-scale structural parameters.}
\label{fig:comparisons_odd}
\end{figure*}

Figure~\ref{fig:comparisons_odd} analyzes the same sets of matched galaxies, but
compares the data to the \fodd{} parameter in GZH to examine its correlation with
irregular morphologies. The relationship in this case is less clear; for the
COSMOS catalogs, in particular, early-types, spirals, and irregulars are
roughly equally probable at \fodd$>0.5$. In both AEGIS and GOODS (there is no
measurment of irregular structure for GEMS in \citealt{hau07}), the proportion
of irregular galaxies does increase with \fodd. However, the actual number of
galaxies at \fodd$>0.5$ is quite small ($N\lesssim100$) in both samples, and so
estimating an accurate fraction is less precise.  We also note that this
particular question ({\it ``anything odd?''}) has somewhat uncertain phrasing
in GZH, and that the design of the tree explicitly separates clumpy from
irregular or merging morphologies, which would have similar properties in
automatic techniques. A more accurate selection of irregular morphologies in
GZH should use a weighted combination of \fbest, \fclumpy, and \fodd.

We repeated all of the analysis comparing morphologies between GZH and external
catalogs on the entire catalog, rather than the volume-limited sample
($z<1.0,m<22.5$) discussed above. If all redshifts and the original magnitude
limit of $m<23.5$ are used, the correlations between all of the automatic
morphological populations and GZH decrease significantly.  The only exception
are the GEMS galaxies, which are the only sample visually classified (similar
to GZH) and had a shallower magnitude cut on their catalog. Based on these
comparisons, we suggest that use of the GZH catalog be limited to galaxies with
$m<22.5$; fainter targets at the sensitivity of the \hst{} Legacy surveys have
been demonstrated to be less reliable, regardless of the method of
morphological classification being used.

@@

We thank Meg Schwamb and the ASIAA for hosting the ``Citizen Science in
Astronomy'' workshop, 3-7 Mar 2014 in Taipei, Taiwan, at which some of this
analysis was done. We thank Jennifer Lotz for sharing her $G$-$M_{20}$
measurements for the AEGIS sample. We thank Coleman Krawczyk for his assistance
in producing Figure~\ref{fig:decisiontree}.

>>>>

% Note: 250-word limit
We present the data release paper for the Galaxy Zoo: Hubble (GZH)
project.\footnote{This publication has been made possible by the participation
of more than 200,000~volunteers in the Galaxy Zoo project. Their contributions
are individually acknowledged at
\url{http://authors.galaxyzoo.org/authors.html}.} This is the third phase in a
large effort to measure reliable, detailed morphologies of galaxies by using
crowdsourced visual classifications of color composite images. Images in GZH
were selected from various publicly-released \textit{Hubble Space Telescope}
Legacy programs using the Advanced Camera for Surveys, with filters that probe
the rest-frame optical emission from galaxies out to $z\sim1$.  The bulk of the
sample is selected to have $m_{I814W}<23.5$, but goes as faint as
$m_{I814W}<26.8$ for deep images combined over 5~epochs. The median redshift of
the combined samples is $\langle z\rangle=0.9\pm0.6$, with a tail extending out
to $z\simeq4$. The GZH morphological data include measurements of both bulge-
and disk-dominated galaxies, details on spiral disk structure that relate to
the Hubble type, bar identification, and numerous measurements of clump
identification and geometry.  This paper also describes a new method for
calibrating morphologies for galaxies of different luminosities and at
different redshifts by using artificially-redshifted galaxy images as a
baseline. The GZH catalog contains both raw and calibrated morphological vote
fractions for $119,849$~galaxies, providing the largest dataset to date
suitable for large-scale studies of galaxy evolution out to $z\sim1$. 
% Number = hubble + goods_shallow + stripe82_coadd
\end{abstract}

@@

\item The Galaxy Evolution from Morphologies and SEDs
    \citep[GEMS;][]{rix04,cal08} survey was centered on the Chandra Deep
    Field-South. The GEMS data covered $\sim800$~arcmin$^2$, completely surrounding
    the area covered by GOODS-S. Images from ACS in GEMS had 1~orbit per pointing
    for a total of 63~pointings. The exposure times were 2160 and 2286~seconds in
    \Vband{} and \zband{}, respectively. The image resolution had a pixel scale
    of 0.03~\arcsec/pixel. The $5\sigma$ limiting magnitude for source
    detection was 25.7~AB in \Vband{} and 24.2~AB in \zband.

@@

@@

The project includes images from the SDSS \citep{yor00,str02}, in particular
those from Data~Release~7 \citep{aba09}. A small number of images were used in
the creation of simulated \hst{} images, described in Sections
\ref{ssec:ferengi} and \ref{sec:debiasing} below; the majority of SDSS images
used in GZH were from Stripe~82. The Stripe~82 samples in GZH include
single-epoch and co-added images. The single-epoch images provide a local
sample for comparison to higher-redshift galaxies (including the new
measurements of clumpy structure), while the co-added images allow for analysis
of morphological properties as a function of image depth.

@@

\begin{figure*}
\center
\includegraphics[width=160mm]{figures/gzh_interface.png}
\caption{Screenshot of the GZH interface (\url{http:/zoo3.galaxyzoo.org}) at
the beginning of classifying a random galaxy, with the classifier ready
to select an answer for the first question in the decision
tree.\label{fig:interface}}
\end{figure*}

@@

Few galaxies in the sample have sufficiently high corrections to completely
change them from being confidently ``smooth'' to ``featured'' following the
correction for redshift bias (Figure~\ref{fig:debiased_corrections}). As a
check, we compare morphologies of highly boosted galaxies to the expert visual
classifications in CANDELS \citep{kar15}.  Only nine galaxies that
were strongly boosted (\ffeatures$<0.25$ and \fbest$>0.50$) in GZH also
appear in the CANDELS expert sample. Of those, \citet{kar15}
classify 4 as spheroids/disky, 2 as disks, 1 as a spheroid, 1 as
irregular/disky and 1 as unclassifiable. The \fbest{} values for GZH are all
between 0.5 and 0.7, making them all intermediate disk candidates. Visual
inspection of the GZH images shows extended disk structure in at least
5~images; the remaining 4 are either extremely faint or have imaging artifacts
in the ACS data.  Since the surveys use different rest-frame filters and the
overlap sample is tiny, though, detailed comparisons between the overall
morphologies are highly difficult.

@@

As with the HST images, each \ferengi{} subject had a varying number of
classifiers answering the various questions in the hierarchical decision tree.
Every classifier answers the first question, {\it ``Is the galaxy smooth and
rounded, with no sign of a disk?''}; as such, the vote fractions \fsmooth,
\ffeatures, and \fartifact{} all have the lowest statistical error for any
question in the tree, based only on the total number of responses (between
40 and 120; see Section~\ref{sec:interface}).  The number of participants
answering subsequent questions, however, is always equal to or less than the
number who answered the preceding question. The average number of responses per
task for fourth- or fifth-tier questions (such as spiral arm structure; Tasks
12--14) is only $4\pm4$ for the \ferengi{} sample. While this distribution is
strongly bimodal (reflecting the true morphologies of selected galaxies), the
very low absolute numbers of votes introduce very high variance when attempting
to calculate a statistical correction.

@@

Another existing study using GZH to select barred galaxies was presented in
\citet{che15}. This study used a slightly different bar sample selection to
demonstrate that AGN hosts show no statistically significant enhancement in bar
fraction or average bar likelihood compared to closely-matched inactive
galaxies. We note that their technique matched test and control galaxies over
the same redshift ranges, which minimizes the possible impact of redshift bias.

@@

COSMOS galaxies have multiple published datasets automatically classifying
morphology, all using a variation of non-parametric measurements. \citet{cas07}
used a combination of concentration ($C$), asymmetry ($A$), $G$, and $M_{20}$
\citep{cas05} make empirical divisions based on several hundred training images
to assign galaxy morphology. A similar method was employed by \citet{tas11},
using the same non-parametric indices but with a different method of
calculating the Petrosian radius and total light profile, and using a
nearest-neighbors method to label morphologies. \citet[][ZEST]{sca07} used $C$,
$A$, $G$, $M_{20}$, the galaxy ellipticity ($\epsilon$), and \sersic{} index
($n$) to quantify morphology; a principal component analysis is used to assign
galaxies to categories. All three methods use discrete morphological categories
of ellipticals, spirals, and peculiar/irregulars. We use the same categories
with the exception of assigning S0/bulge-dominated disks (type~2.0) from ZEST
into the early-type category.

@@

Galaxies from the three COSMOS morphology catalogs all have morphologies that are
reasonably well-predicted by the GZH based on \fbest. The relation is strongest
for early-type galaxies, since the comparable fractions of irregulars and
spirals shows that clumpy/asymmetric morphologies are clearly also being
captured by the \fbest{} parameter. The small dropoff in the fraction of
early-types for the Cassata and ZEST samples at \fbest$<0.1$ is somewhat
puzzling, but could potentially be due to the presence of a small
population of extremely compact objects (stars or quasars) that would be
classified as artifacts in GZH, explaining the dip in \fbest.

%We also emphasize that the previously published COSMOS catalogs are not
%internally consistent; for example, only 64\% of the galaxies in \citet{tas11}
%have the same morphological class as \citet{cas07}. The bulk of these are
%galaxies classified as ``spiral'' using one technique and ``irregular'' using
%another (\citealt{cas07} is more likely to classify galaxies as irregular).  A
%detailed comparison of the external catalogs is outside the scope of this
%paper, but we emphasize only that it is \emph{impossible} for any catalog to be
%consistent with all three.  We interpret the broad agreements between
%early-types, spirals, and irregulars for all three samples as a validation of
%the crowdsourced visual classifications provided by GZH.

%\begin{figure*}
%\center
%\includegraphics[width=1.0\textwidth]{figures/comparisons_odd.pdf}
%\caption{Same as Figure~\ref{fig:comparisons_features}, but comparing external
%morphologies to \fodd{} in GZH. Measurement of \fodd{} excludes galaxies
%labeled as ``odd'' due to lenses or dust lanes ($<1\%$ of the total), which are GZH
%categories not expected to correlate with larger-scale structural parameters.}
%\label{fig:comparisons_odd}
%\end{figure*}
%
%Figure~\ref{fig:comparisons_odd} analyzes the same sets of matched galaxies, but
%compares the data to the \fodd{} parameter in GZH to examine its correlation with
%irregular morphologies. The relationship in this case is less clear; for the
%COSMOS catalogs, in particular, early-types, spirals, and irregulars are
%roughly equally probable at \fodd$>0.5$. In both AEGIS and GOODS (there is no
%measurement of irregular structure for GEMS in \citealt{hau07}), the proportion
%of irregular galaxies does increase with \fodd. However, the actual number of
%galaxies at \fodd$>0.5$ is quite small ($N\lesssim100$) in both samples, and so
%estimating an accurate fraction is less precise.  We also note that this
%particular question ({\it ``anything odd?''}) has somewhat uncertain phrasing
%in GZH, and that the design of the tree explicitly separates clumpy from
%irregular or merging morphologies, which would have similar properties in
%automatic techniques. A more accurate selection of irregular morphologies in
%GZH should use a weighted combination of \fbest, \fclumpy, and \fodd.

A detailed comparison of the external catalogs is outside the scope of this
paper, but we emphasize only that it is impossible for any catalog to be fully
consistent with all of the previously published datasets (for example, only
64\% of the galaxies in \citealt{tas11} have the same morphological class as
the same galaxies in \citealt{cas07}). We interpret the broad agreements
between early-types, spirals, and irregulars between GZH and all three methods
of automated classification as a validation of both the crowdsourced visual
classifications provided by GZH and those automated methods.

It has been argued that automated methods are preferable to visual
classification due to being quantified and more easily reproducible
\citep[\eg,][]{con14,paw16}. We argue that this distinction applies only to
visual classification done by individual and/or small numbers of experts. The
technique of crowdsourcing using a large number of independent classifiers
provides visual classification which is both quantified, and reproducible, and
(as demonstrated in this section), broadly agrees with all the automated
methods and expert visual classifications.

We repeated all analyses comparing morphologies between GZH and external
catalogs on the entire catalog, rather than the volume-limited sample
($z<1.0,m<22.5$) discussed above. If galaxies at all redshifts and at the
original magnitude limit of $m<23.5$ are used, the correlations between all of
the automatic morphological populations and GZH decrease significantly.  The
exceptions are the GEMS galaxies, which are the only sample that was visually
classified and had a shallower magnitude cut on their catalog.
Based on these comparisons, we suggest that use of the GZH catalog be limited
to galaxies with $m_\mathrm{I|i|z}<22.5$; fainter targets at the sensitivity of the \hst{}
Legacy surveys have been demonstrated to be less reliable, regardless of the
method of morphological classification being used.

@@

We thank Meg Schwamb and the ASIAA for hosting the ``Citizen Science in
Astronomy'' workshop, 3-7 Mar 2014 in Taipei, Taiwan, at which some of this
analysis was initiated. We thank Jennifer Lotz for sharing her $G$-$M_{20}$
measurements for the AEGIS sample. We thank Coleman Krawczyk for his assistance
in producing Figure~\ref{fig:decisiontree}.
